Red States = Red Flags?

### 2024-02-15 paper/paper.qmd
Add political affiliation and education to paper

# Workspace Set up
library(tidyverse)
library(haven)
library(tidyr)
library(readxl)
library(dplyr)
library(readr)
library(knitr)
library(ggplot2)
```

@@

Education

@@

>>>>

### Preparing Workspace
# Install and load necessary packages
if (!requireNamespace("tidyverse", quietly = TRUE)) install.packages("tidyverse")
if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")
if (!requireNamespace("tidyr", quietly = TRUE)) install.packages("tidyr")
if (!requireNamespace("ggplot2", quietly = TRUE)) install.packages("ggplot2")
if (!requireNamespace("knitr", quietly = TRUE)) install.packages("knitr")
if (!requireNamespace("readr", quietly = TRUE)) install.packages("readr")
if (!requireNamespace("maps", quietly = TRUE)) install.packages("maps")
if (!requireNamespace("haven", quietly = TRUE)) install.packages("haven")
if (!requireNamespace("readxl", quietly = TRUE)) install.packages("readxl")
if (!requireNamespace("here", quietly = TRUE)) install.packages("here")

# Workspace Set up
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(readr)
library(maps)
library(haven)
library(readxl)
library(here)
```

@@

@@

# Education
```{r echo=FALSE, message=FALSE}
#Education

# Load Education data
educ_data <- read_csv("../data/analysis_data/educ/numbirths_educ_2044.csv")

# Creating the plot
ggplot(educ_data, aes(x = year)) +
  geom_line(aes(y = numbirth_lesshs, color = "Less than HS"), linewidth = 1) +
  geom_line(aes(y = numbirth_hsgrad, color = "HS Graduate"), linewidth = 1) +
  geom_line(aes(y = numbirth_somecoll, color = "Some College"), linewidth = 1) +
  geom_line(aes(y = numbirth_colgrad, color = "College Graduate"), linewidth = 1) +
  geom_line(
    aes(y = numbirthalleduc, color = "All Education Levels"),
    size = 1,
    linetype = "dashed"
  ) +
  scale_color_manual(
    values = c(
      "Less than HS" = "blue",
      "HS Graduate" = "green",
      "Some College" = "orange",
      "College Graduate" = "red",
      "All Education Levels" = "grey"
      )
  ) +
  labs(
    title = "Number of Births by Education Level Over Time",
    x = "Year",
    y = "Number of Births",
    color = "Education Level"
  ) +
  theme_minimal() +
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

# Display the plot
ggsave("births_by_education_level_over_time.png",
       width = 10,
       height = 6)

```
# State Political Affiliation and Birth Rates
```{r echo=FALSE, message=FALSE}
# https://stackoverflow.com/questions/29381652/plotting-us-heat-map-with-text

#reading the UKNCPR Datat
poldata <- read_csv(here("data", "analysis_data", "UKCPR_National_Welfare_Data_Update_020623.csv"))

poldata <- poldata %>%
  select(state_name,
         year,
         Population,
         `Fraction of State Senate that is Democrat`)

# Create a mapping of state abbreviations to full names
state_abbrevs <- state.abb
state_names <- state.name

# Create a data frame for the mapping
state_mapping <-
  data.frame(abbrev = tolower(state_abbrevs),
             name = tolower(state_names))

#Get numbirths data
numbirths_data <- read_csv(here("data", "analysis_data", "annual_policy", "numbirths_2001_2019.csv"))
births_2019 <- numbirths_data %>% filter(year == 2019)

# Convert state abbreviations to full names in your datasets
numbirths_2019_fullname <- births_2019 %>%
  mutate(stname = tolower(stname)) %>%
  left_join(state_mapping, by = c("stname" = "abbrev")) %>%
  select(-stname) %>%
  rename(stname = name)

poldata_2019 <- poldata %>% filter(year == 2019)

poldata_2019_fullname <- poldata_2019 %>%
  mutate(state_name = tolower(state_name)) %>%
  left_join(state_mapping, by = c("state_name" = "abbrev")) %>%
  select(-state_name) %>%
  rename(state_name = name)

# Merge poldata with numbirths_data using full state names
data_merged <-
  left_join(numbirths_2019_fullname,
            poldata_2019_fullname,
            by = c("stname" = "state_name"))

states_map <- map_data("state")

# Merge with states map data
map_data_merged <-
  left_join(states_map, data_merged, by = c("region" = "stname"))
```

```{r echo=FALSE, message=FALSE}
# Plot
ggplot(map_data_merged) +
  geom_polygon(
    aes(
      x = long,
      y = lat,
      group = group,
      fill = `Fraction of State Senate that is Democrat`
    ),
    color = "white"
  ) +
  geom_text(
    data = map_data_merged %>% group_by(region) %>% summarise(
      long = mean(long),
      lat = mean(lat),
      numbirths = first(births)
    ),
    aes(x = long, y = lat, label = numbirths),
    size = 3,
    check_overlap = TRUE
  ) +
  scale_fill_gradient(low = "red",
                      high = "blue",
                      name = "Democrat Fraction") +
  coord_fixed(1.3) +
  labs(title = "2019 Births and Democrat Fraction in State Senate by State") +
  theme_void() +
  theme(legend.position = "right")
```
